### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
Lots of changes on Exp3, Exp4 etc notations
Still changing the macros

Following this Definition~\ref{def:2:KLDivergence} for two distributions
$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ in the same exponential family,
we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) \eqdef \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$,
%
% \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
$\KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2)$.
% \end{equation}
%
Without diving more in the details of exponential families,
it is interesting to illustrate this definition and the notations with two important examples:

\label{par:2:notationsExponentialFamiliesBernoulliGaussian}
\begin{itemize}
    \item
    % $\bullet$
    \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
    and for a Bernoulli distribution of mean $\mu\in[0,1]$, denoted $B(\mu)$,
    the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \mathrm{e}^{\eta})$
    (with the limit behavior $\eta=+\infty$ if $\mu=1$).

@@

The different algorithms being compared are the following, in the order they were presented above:
%
first we include the pure-exploration and
pure-exploitation algorithms,
then the $\varepsilon$-greedy (with $\varepsilon_t=\varepsilon_0/t$ and $\varepsilon_0=6K/d^2=3000$) and
Explore-then-Commit (with $T_0=K4/d^2 \log(d^2T/4)=2000\log(T/400)$) algorithms
from Section~\ref{sub:2:naiveSimpleStrategies}, which both uses a tuning based on a prior knowledge of $\Delta$ (that gives large values of $\varepsilon_0$ and $T_0$ yielding linear regret for small horizon).
%
We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (their are robust and efficient on unknown problems):
$\UCB_1$ (with $\alpha=1$),
\klUCB{} (with Bernoulli $\kl = \KL_{B}$),
and Thompson sampling (with Beta posteriors and flat uniform priors).
%

@@

In this chapter, we presented the multi-armed bandit model, focussing on a finite number of arms, and real-valued rewards.
Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
By showcasing an interactive demonstration made available online,
% XXX ?
% (\href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}),
we presented the notations used in all this thesis.
%
% The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
% % and corresponds to our article \cite{Besson2018WCNC}.
% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
% Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
% We proposed an extension of the Exp4 algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
% %
We then used the small Bernoulli-distributed problem behind this demonstration
to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

>>>>

Following this Definition~\ref{def:2:KLDivergence} for two distributions
$d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ in the same exponential family $\cE$,
we can use a shorter notation and write $\KL( \eta_1, \eta_2 ) \eqdef \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, which can be simplified to use only the two parameters $\eta_1$ and $\eta_2$ of $d_1$ and $d_2$,
and $\mu_1 = \E_{\eta_1}[X]$ the mean of distribution $d_1$,
%
% \begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
$\KL_{\cE}\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2) X ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2) \mu_1 - A(\eta_1) + A(\eta_2)$.
% \end{equation}
%
Without diving more in the details of exponential families,
it is interesting to illustrate this definition and the notations with two important examples:

\label{par:2:notationsExponentialFamiliesBernoulliGaussian}
\begin{itemize}
    \item
    % $\bullet$
    \textbf{Bernoulli distributions} can be seen as an exponential family with $h(x) = 1$,
    and for a Bernoulli distribution of mean $\mu\in[0,1]$, denoted $B(\mu)$,
    the parameter is $\eta = \mu / (1 - \mu)$, giving $A(\eta) = \log(1 + \e^{\eta})$
    (with the limit behavior $\eta=+\infty$ if $\mu=1$).

@@

The different algorithms being compared are the following, in the order they were presented above:
%
first we include the pure-exploration and
pure-exploitation algorithms,
then the $\varepsilon$-greedy (with $\varepsilon_t=\varepsilon_0/t$ and $\varepsilon_0=6K/d^2=3000$) and
Explore-then-Commit (with $T_0=K4/d^2 \log(d^2T/4)=2000\log(T/400)$) algorithms
from Section~\ref{sub:2:naiveSimpleStrategies}, which both uses a tuning based on a prior knowledge of $\Delta$ (that gives large values of $\varepsilon_0$ and $T_0$ yielding linear regret for small horizon).
%
We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (their are robust and efficient on unknown problems):
$\UCB_1$ (with $\alpha=1$),
\klUCB{} (with Bernoulli $\kl = \KL_{\cB}$),
and Thompson sampling (with Beta posteriors and flat uniform priors).
%

@@

In this chapter, we presented the multi-armed bandit model, focussing on a finite number of arms, and real-valued rewards.
Our main focus is on one-dimensional exponential families of distributions, and on stochastic and stationary problems.
By showcasing an interactive demonstration made available online,
% XXX ?
% (\href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}),
we presented the notations used in all this thesis.
%
% The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
% % and corresponds to our article \cite{Besson2018WCNC}.
% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
% Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
% We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
% %
We then used the small Bernoulli-distributed problem behind this demonstration
to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

### 2019-07-04 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex
Typos, et commentaires Christophe
Still changing the macros

% \subsection{Proof of Proposition \ref{prop:41:Lagrangian}}
% \label{sec:5:proofLagrangian}

@@

    \item
    So the gradient is zero $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
    iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\mathrm{e}^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
    and with $y = 1 + x$, we get $\mathrm{e}^{y}y = \lambda\mathrm{e} / (1-p)^{S_i - 1}$.

	\item
    By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \mathrm{e}^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1$.
    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, x/\log(1-p)) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\mathrm{e} / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
    This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,
    and so the unique solution to the maximization problem \eqref{eq:41:optPb}, thanks to \cite[Theorem 1]{Luenberger68}.
    \hfill{}$\square$
\end{itemize}

>>>>

% \subsection{Proof of Proposition~\ref{prop:41:Lagrangian}}
% \label{sec:5:proofLagrangian}

@@

    \item
    So the gradient is zero $\frac{\partial}{\partial D_i} \mathcal{L}(\boldsymbol{D}, \lambda) |_{D_i=D_i^*} = 0$
    iff $D_i^*$ satisfies $(1-p)^{D_i^*}( 1 + \log(1-p) D_i^*) = \lambda / (1-p)^{S_i - 1}$. Let $x = \log(1-p) D_i^*$ this is equivalent to $\e^{x}(1 + x) = \lambda / (1-p)^{S_i - 1}$
    and with $y = 1 + x$, we get $\e^{y}y = \lambda\e / (1-p)^{S_i - 1}$.

	\item
    By using the $\mathcal{W}$-Lambert function $\mathcal{W}$ \cite{Corless96}, reciprocal of $y \mapsto \e^{y}y$, we get $x = y - 1 = \mathcal{W}(\lambda\e / (1-p)^{S_i - 1}) - 1$.
    So the gradient of the Lagrangian is zero iff $D_i^* = \max(0, x/\log(1-p)) = \left[ \frac{1}{\log(1-p)}\mathcal{W}(\lambda\e / (1-p)^{S_i - 1}) - 1\right]^{+}$, because $D_i^*$ has to be non-negative.
    This gives the $i$-th coordinate of the unique saddle point of $f(\boldsymbol{D})$,
    and so the unique solution to the maximization problem \eqref{eq:41:optPb}, thanks to \cite[Theorem 1]{Luenberger68}.
    \hfill{}$\square$
\end{itemize}
